Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-14 surfaces-write-up/surface-write-up.tex
removed stupid ;
what? I made changes since the last push?

\begin{prop}
\label{prop:cone-generation}
Let $n,t \in \bz,$ let $\alpha_1, \ldots, \alpha_n \in \bq$ and let $k_i^j \in \bz$ with $1 \leq i \leq n, 1 \leq j \leq t.$ Define
\begin{align*}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}
\end{align*}
Let $e_1, \ldots, e_t \in \Sigma,$ with $e_i = (\pdeg_i, c_1^i, \ldots, c_n^i)$ be a set of {\bf extremal rays} of $\Sigma$, meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of $e_1, \ldots, e_t$.  Then, as a semigroup, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\pdeg_i$. Furthermore, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{t} w_i e_i,$ with the first coordinate of $\lambda$ less than $\sum_{i=1}^{t}\pdeg_i$ and $w_1,\ldots, w_t \in \bz$.

@@

\begin{lem}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^n$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$.  Then $R_D$ is generated in degree at most    and relations in degree at most  . \todo{insert numbers}
\end{lem}

>>>>

\begin{prop}
\label{prop:cone-generation}
Let $n,t \in \bz,$ let $\alpha_1, \ldots, \alpha_n \in \bq$ and let $k_i^j \in \bz$ with $1 \leq i \leq n, 1 \leq j \leq t.$ Define
\begin{align*}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}
\end{align*}
Let $e_1, \ldots, e_r \in \Sigma,$ with $e_i = (\pdeg_i, c_1^i, \ldots, c_n^i)$ be a set of {\bf extremal rays} of $\Sigma$, meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of $e_1, \ldots, e_r$.  Then, as a semigroup, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\pdeg_i$. Furthermore, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{t} w_i e_i,$ with the first coordinate of $\lambda$ less than $\sum_{i=1}^{t}\pdeg_i$ and $w_1,\ldots, w_t \in \bz$.

@@

\begin{lem}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$.  Then $R_D$ is generated in degree at most $\sum_{i=1}^n \ell_i a_i$.
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align*}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}.
\end{align*}
We note that $\Sigma$ has extremal rays given by the lattice points $e_i$ corresponding to
\[
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}
\]
for each $i\in \{1, \ldots n\}$.  Therefore, applying proposition \ref{prop:cone-generation}, we see $R_D$ is generated in degree less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}
